Development of a educational and sustainable robotic platform for children and adults

### 2013-12-17 doc/paper.tex
foo

\begin{table}[h!]
\centering
\begin{tabular}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{4}    & 99 CHF (without taxes directly from the Ecole Polytechnique Fédérale de Lausanne) is a reasonable price for a robot with this huge amount of features. On retail stores the price is higher due to profit margin.\\
Educational & \stars{3}     & Thymio II does not yet have a repository of educational material only, but it is planned to release such material. The compability with LEGO is a huge bonus, and creates a lot of pplications on which educational problems can be explained and demonstrated.\\
Sustainable  & \stars{3}     & The robot consists of small SMD parts and therefore parts cannot be replaced by the target user group. Materials used are suitable for children and designed to last. The integrated Li-Po battery can be recharged and therefore produces less environmental waste.\\
Extendable & \stars{3} & Except the LEGO compatible fixation points, Thymio II is not extendable without unsoldering components. On the software side, the choice of microcontroller supports a lot of additional program features.\\
Trouble-free & \stars{4} & Thymio II is trouble free since it comes preassembled and the usage is straight forward.\\
Open-Source & \stars{5} & Everything needed to build the robot is licensed under a permissive creative commons license. Program files and libraries are licensed under a LGPL license. Even the CAD files for the plastic mold injection parts are freely available.\\
\bottomrule
\end{tabular}
\caption{Thymio II evaluation}
\label{tbl:thymio_eval}
\end{table}

@@

In terms of cost effectiveness a solution with two wheels powered by a DC motor each performs well. A DC motor with integrated gearbox including (see figure @TODO) is available for 5.5 USD, the plastic wheel costs 3.5, a motor driver such as the L293 costs about 3.5 USD rendering a total cost of 21.5 USD per robot. In addition speed sensors for each of the wheels are needed since the motors do not have exactly the same speed. Without speed control the robot cannot drive straight forward.

@@

\subsubsection{Wheeled with two Stepper motors}
Stepper motors are slightly modified DC motors. In stepper motors the full rotation is divided into a number of equal steps which can be separately controlled. The motor can step forward, backward or hold the current position with huge torque. Stepper motors are usually shipped within a gear box which reduces steps much further to gain higher precision and torque. 
In typical applications stepper motors are driven by a dual H-bridge or darlington arrays attached to shift registers. The latter method is very cost effective but limited in current supply, the first method utilized a motor driver such as the L293 per stepper.

Stepper motors are availble for about 5 USD (in the USA\footnote{http://www.adafruit.com/products/858}). Two stepper motors are needed to drive the robot, an additional ball caster is needed to stabilize the movement. As mentioned in the forehand the cost of driving a stepper motor can be reduced using a shift register such as the 74HC595 in combination with a darlinton array such as the ULN2803. Costs for this solution are about 12 USD where the stepper motors cost 10 USD, and the integrated circuits about 2 USD.

@@

\subsubsection{Conclusion}
All in all a solution using stepper motors with darlington array connected to shift registers is the most cost-effective solution. It provides the necessary functions and complies to the evaluation aspects such as the educational potential.

\subsection{Sensors}
Sensors are important to make the robot sense its environment. Sensing the environment enables the interaction in it, the avoidance of obstacles or

\subsection{IR Reciever}

\subsection{IR Proximity Sensor}

\subsection{Ultrasonic Distance Sensor}

\subsection{Accelerometer}

@@

\subsection{Compass}

\subsubsection{Conclusion}
\subsection{Communication}
\subsubsection{Infrared}
\subsubsection{RFM12}
\subsubsection{Bluetooth}
\subsubsection{NRF24L01}

>>>>

\begin{table}[h!]
\centering
\begin{tabular}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{4}    & 99 CHF (without taxes directly from the Ecole Polytechnique Fédérale de Lausanne) is a reasonable price for a robot with this huge amount of features. On retail stores the price is higher due to profit margin.\\
Educational & \stars{3}     & Thymio II does not yet have a repository of educational material only, but it is planned to release such material. The compability with LEGO is a huge bonus, and creates a lot of pplications on which educational problems can be explained and demonstrated.\\
Sustainable  & \stars{3}     & The robot consists of small SMD parts and therefore parts cannot be replaced by the target user group. Materials used are suitable for children and designed to last. The integrated Li-Po battery can be recharged and therefore produces less environmental waste.\\
Extendable & \stars{3} & Except the LEGO compatible fixation points, Thymio II is not extendable without unsoldering components. On the software side, the choice of microcontroller supports a lot of additional program features.\\
Trouble-free & \stars{4} & Thymio II is trouble free since it comes preassembled and the usage is straight forward.\\
Open-Source & \stars{5} & Everything needed to build the robot is licensed under a permissive creative commons license. Program files and libraries are licensed under a LGPL license. Even the CAD files for the plastic mold injection parts are available without a charge.\\
\bottomrule
\end{tabular}
\caption{Thymio II evaluation}
\label{tbl:thymio_eval}
\end{table}

@@

In terms of cost effectiveness a solution with two wheels powered by a DC motor each performs well. A DC motor with integrated gearbox (see figure @TODO) is available for 5.5 USD, the plastic wheel costs 3.5 USD, a motor driver such as the L293 costs about 3.5 USD rendering a total cost of 21.5 USD per robot. In addition speed sensors for each of the wheels are needed since the motors do not have exactly the same speed. Without speed control the robot cannot drive straight forward.

@@

\subsubsection{Wheeled with two Stepper motors}
Stepper motors are basically DC motors with more coils. In stepper motors the full rotation is divided into a number of equal steps which can be separately controlled. The motor can step forward, backward or hold the current position with huge torque. Stepper motors are usually shipped within a gear box which reduces steps much further to gain higher precision and torque. 
In typical applications stepper motors are driven by a dual H-bridge or darlington arrays attached to shift registers. The latter method is very cost effective but limited in current supply, the first method utilized a motor driver such as the L293 per stepper.

Stepper motors are availble for about 5 USD (in the USA\footnote{http://www.adafruit.com/products/858}). Two stepper motors are needed to drive the robot, an additional ball caster is needed to stabilize the movement. As mentioned in the forehand the cost of driving a stepper motor can be reduced using a shift register such as the 74HC595 in combination with a darlington array such as the ULN2803. Costs for this solution are about 12 USD where the stepper motors cost 10 USD, and the integrated circuits about 2 USD.

@@

\subsubsection{Conclusion}
All in all a solution using stepper motors with darlington array connected to shift registers is the most cost-effective solution. It provides the necessary functions and complies to the evaluation aspects such as the educational potential. Its usage is rather trouble free when apropriate libraries for motor control are provided.

\subsection{Sensors}
Sensors are important to make the robot sense its environment. Sensing the environment enables interaction such as obstacle avoidance, finding objects or following lines. Connecting sensors to a microcontroller can be done in various ways using various protocols. Different protocols require a different amount of microcontroller pins. Dependening on the sensor technology more or less computing time is needed to evaluate the sensors value.

\subsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

IR recievers are widely used in remote control applications such as televisions and remote controlled toys. Sending a signal requires a LED which emits light in the infrared spectrum. Cheap multipurpose remotes (about 1 USD) can be used to control own applications. Infrared beacons which steadily send a signal can be used to give a robot information on its current location.

Attaching a reciever to a microcontroller requires a single input pin which can be either a digital pin or a pin internally attached to an ADC. In terms of cost effectiveness and education potential IR recievers are a must have sensor. Under the premise that library for infrared readings exist, the sensor is easy to use and trouble free.

\subsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29 m / s at sea level).

The price for an ultrasonic distance sensor is very low, cheap versions with a range between 1 and 255 centimeters start at about 1.5 USD. Interfacing the sensor requires two microcontroller pins - one for triggering the echo and another one for reading back the distance value. The trigger pin can be a generic purpose input/output pin, the sense pin must be attached to a pin which provides an analog to digital converter.

A typical application for ultrasonic distance sensor is obstacle avoidance. By mounting the sensor on a servo motor the visible area can be increased which allows better navigation due to enhanced obstacle avoidance possibilities.

\subsection{Accelerometer}
Accelerometers can sense acceleration on one or more axis. This can be used to determine the speed of the robot

@@

\subsection{Compass}
A compass allows the robot

\subsubsection{Conclusion}
\subsection{Communication}
Communication between robots or between the robot is essential to widen the application possibilities. When the robot is connected to a computer it gets much more processing power to solve complicated problems; when robots can communicate with each other tasks can be solved collaboratively.

\subsubsection{Infrared}
Infrared based data transfer is widely used in remote control application where a small amount of data is transferred. Infrared communication is easily interfered by other infrared signals. This limits the data rate since error correction mechanisms must be implemented.

When using infrared signals only one robot can send at a time which requires collision detection. The hardware requirements for infrared transmissions are ver low, only a infrared reciever and a infrared LED is required which sums up to a total of 0.5 USD. The infrared reciever and the infrared LED require one free GPIO pin on the microcontroller each.

Since a lot of effort is needed to make the connection stable, this solution is not trouble free, although techniques to stabilize communication have educational potential.

\subsubsection{RFM12}
\subsubsection{Bluetooth}
\subsubsection{NRF24L01}
